Oral anticoagulants

### 2021-01-15 r_notebooks/exploring_oral_anticoagulants_war_prop.Rmd
r analyses

```{r setup}
## Packages
devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(serCymruTools)

@@

@@

## Function

@@

# Modifying functions to model log(n) as using proportion data
fableModels <- function(data){
  if(!is_tsibble(data)){data <- as_tsibble(data, index = datename)}

  data %>%
    mutate(datename = yearmonth(datename)) %>%
    model(trend_model1 = TSLM(log(n) ~ trend()),#time series linear model
          trend_model2 = TSLM(log(n) ~ trend() + season("year")),
          ets1 = ETS(log(n) ~ trend()),
          ets2 = ETS(log(n) ~ trend() + season("A")),
          ets3 = ETS(log(n) ~ trend() + season("M")),
          arima = ARIMA(log(n), stepwise = FALSE, approximation = FALSE),
          comb1 = combination_model(TSLM(log(n) ~ trend()),
                                    ETS(log(n) ~ trend())),
          comb2 = combination_model(TSLM(log(n) ~ trend() + season("year")),
                                    ETS(log(n) ~ trend() + season("A"))),
          stl_dcmp1 = decomposition_model(STL(log(n) ~ trend(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          stl_dcmp2 = decomposition_model(STL(log(n) ~ trend() + season(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          s_naive = SNAIVE(log(n)))
}

@@

  oc_prophet_model1 <-
    fable.prophet::prophet(log(n) ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model2 <-
    fable.prophet::prophet(log(n) ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model3 <-
    fable.prophet::prophet(log(n) ~ growth("linear") +
                             season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model4 <-
    fable.prophet::prophet(log(n) ~ growth("linear") +
                             season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model5 <-
    fable.prophet::prophet(log(n) ~ season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model6 <-
    fable.prophet::prophet(log(n) ~ season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model7 <-
    fable.prophet::prophet(log(n) ~ holiday(holidays))
  oc_prophet_model8 <-
    fable.prophet::prophet(log(n))

@@

            tibble(holiday = "easter", # Sunday
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -2,
                   upper_window = 1))
regionalCV <- function(data, cv_dist = 6, init = 36, step = 1){

@@

  fableCV <- function(data){ # Return CV model accuracy for fableModels()
    data_test <-
      data %>%
      mutate(datename = yearmonth(datename))

@@

    data_tr %>%
      fableModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)
  }

  prophetCV <- function(data){# Return CV model accuracy for prophetModels()
    data_test <-
      data %>%
      mutate(datename = yearmonth(datename))

    data_tr %>%
      prophetModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)
  }

  plan(cluster, 8) # Set up parallel processing to decrease computation time
  #plan(multisession) # Set up parallel processing to decrease computation time
  # Prophet models won't run with plan(multisession) so need to be run
  # sequentially. Is it worth trying to write this in Python and pass it back
  # to R?

  # Probably worth trying the prophet::cross_validation() method. If it
  # returns RMSE and is faster it will be useful.
  prophet_fits <-
    future_map_dfr(.x =  regions,
                   .f = ~prophetCV(data = filter(data, regional_unit == .x)),
                   .options = furrr_options(seed = TRUE))
  message("prophetModels() complete")

  # Fable models will run using plan(multisession), but need to be run as below
  # with `furrr_options(seed = TRUE)` otherwise improper random numbers are
  # generated
  fable_fits <-
    future_map_dfr(.x =  regions,
                   .f = ~fableCV(data = filter(data, regional_unit == .x)),
                   .options = furrr_options(seed = TRUE))
  plan(sequential) # Close the multiprocess function to prevent memory leakage
  message("fableModels() complete")

@@

regionalJointFcsts <- function(data, h ="5 months"){
  regions <- unique(data$regional_unit)
  plan(cluster)
  fable_fit <-
    future_map_dfr(.x = regions,
                   .f = ~fableModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("fable fits complete")

  fable_fcst <- fable_fit %>%
    forecast(h = h) %>%
    as_tibble()
  message("fable forecasts complete")

  prophet_fit <-
    future_map_dfr(.x = regions,
                   .f = ~prophetModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("prophet fits complete")

  prophet_fcst <-
    forecast(prophet_fit, h = h) %>%
    as_tibble()
  # Some prophet forecasts are giving daily forecasts. We want monthly forecasts.
  # This may also be influencing accuracy

  message("prophet forecasts complete")
  plan(sequential)
  # prophetModels now being run in parallel. Needed to reinstall prophet from
  # source

@@

# Running regionalCV() on each of the quarter of the data
system.time(oac_p_gp_cv_models_1 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_1 %>%
              regionalCV())
system.time(oac_p_gp_cv_models_2 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_2 %>%
              regionalCV())
system.time(oac_p_gp_cv_models_3 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_3 %>%
              regionalCV())
system.time(oac_p_gp_cv_models_4 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_4 %>%
              regionalCV())

@@

oac_p_gp_cv_models %>%
  write_rds("../data/oac_p_gp_cv_models.rds")

@@

```{r}
# regionalJointFcsts <- function (data, h = "5 months", keep_fits = FALSE) 
# {
#   if (!is.logical(keep_fits)) {
#     stop("`keep_fits` must be of type logical")
#   }
#   regions <- unique(data$regional_unit)
#   
#   cl <- parallel::makeCluster(8)
#   plan(cluster, workers = cl)
#   
#   fable_fcst <- 
#     future_map_dfr(.x = regions, 
#                    .f = ~fableModels(filter(data, 
#                                             regional_unit == .x)) %>% 
#                      forecast(h = h) %>% as_tibble(), 
#                    .options = furrr_options(seed = TRUE))
#   
#   
#   message("fable fits complete, step 1/4")
#   message("fable forecasts complete, step 2/4")
# 
#   prophet_fcst <- 
#     future_map_dfr(.x = regions, 
#                    .f = ~prophetModels(filter(data, 
#                                             regional_unit == .x)) %>% 
#                      forecast(h = h) %>% as_tibble(), 
#                    .options = furrr_options(seed = TRUE))
#   message("prophet fits complete, step 3/4")
#   message("prophet forecasts complete, step 4/4")
#   
#   parallel::stopCluster(cl)
#   plan(sequential)
#   joint_fcst <- dplyr::bind_rows(fable_fcst, prophet_fcst) %>% 
#     dplyr::mutate(lb_95 = quantile(n, 0.025), lb_80 = quantile(n, 
#       0.1), med = median(n), ub_80 = quantile(n, 0.9), 
#       ub_95 = quantile(n, 0.975), sd = distributional::variance(n)^2, 
#       dist = n, n = NULL)
#       return(list(forecasts = joint_fcst))
# }

system.time(oac_p_gp_fcsts <-
              oac_p_monthly_quantity_gp_pre_cv19 %>%
              filter(!regional_unit %in% incomplete_gp) %>%
              regionalJointFcsts())

oac_p_gp_fcsts %>%
  write_rds("../data/oac_p_gp_fcsts.rds")

@@

>>>>

```{r setup}
## Packages
devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(progressr)
library(beepr)
library(serCymruTools)

@@

handlers(global = TRUE)
handlers(
  list(
    handler_progress(
      format = ":spin :current/:total (:message) [:bar] :percent in :elapsed ETA: :eta",
      width = 60,
      complete = "+"),
    handler_beepr(finish   = "wilhelm",
                  interval = 2.0)
  )
)

@@

## Function
# Transforming data to be bound beweent a lower and upper limit http://www.maths.bristol.ac.uk/R/web/packages/fable/vignettes/transformations.html
scaled_logit <- function(x, lower=0, upper=1){
  log((x-lower)/(upper-x))
}
inv_scaled_logit <- function(x, lower=0, upper=1){
  (upper-lower)*exp(x)/(1+exp(x)) + lower
}
my_scaled_logit <- fabletools::new_transformation(scaled_logit, inv_scaled_logit)

@@

# Modifying functions to model my_scaled_logit(n, 0, 100) as using proportion data
fableModels <- function(data){
  if(!is_tsibble(data)){data <- as_tsibble(data, index = datename)}

  data %>%
    mutate(datename = yearmonth(datename)) %>%
    model(trend_model1 = TSLM(my_scaled_logit(n, 0, 100) ~ trend()),#time series linear model
          trend_model2 = TSLM(my_scaled_logit(n, 0, 100) ~ trend() + season("year")),
          ets1 = ETS(my_scaled_logit(n, 0, 100) ~ trend()),
          ets2 = ETS(my_scaled_logit(n, 0, 100) ~ trend() + season("A")),
          ets3 = ETS(my_scaled_logit(n, 0, 100) ~ trend() + season("M"), restrict = FALSE),
          arima = ARIMA(my_scaled_logit(n, 0, 100), stepwise = FALSE, approximation = FALSE),
          comb1 = combination_model(TSLM(my_scaled_logit(n, 0, 100) ~ trend()),
                                    ETS(my_scaled_logit(n, 0, 100) ~ trend())),
          comb2 = combination_model(TSLM(my_scaled_logit(n, 0, 100) ~ trend() + season("year")),
                                    ETS(my_scaled_logit(n, 0, 100) ~ trend() + season("A"))),
          stl_dcmp1 = decomposition_model(STL(my_scaled_logit(n, 0, 100) ~ trend(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          stl_dcmp2 = decomposition_model(STL(my_scaled_logit(n, 0, 100) ~ trend() + season(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          s_naive = SNAIVE(my_scaled_logit(n, 0, 100)))
}

            tibble(holiday = "easter", # Sunday
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -2,
                   upper_window = 1))

@@

  oc_prophet_model1 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model2 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model3 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ growth("linear") +
                             season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model4 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ growth("linear") +
                             season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model5 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ season("year", type = "additive") +
                             holiday(holidays))
  oc_prophet_model6 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ season("year", type = "multiplicative") +
                             holiday(holidays))
  oc_prophet_model7 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100) ~ holiday(holidays))
  oc_prophet_model8 <-
    fable.prophet::prophet(my_scaled_logit(n, 0, 100))

@@

regionalCV <- function(data, cv_dist = 6, init = 36, step = 1){

@@

  p <- progressr::progressor(along = regions)

  fableCV <- function(data){ # Return CV model accuracy for fableModels()
    x <- data$regional_unit

    data_test <-
      data %>%
      mutate(datename = yearmonth(datename))

@@

    data_tr %>%
      fableModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)

    p()
  }

  prophetCV <- function(data){# Return CV model accuracy for prophetModels()
    x <- data$regional_unit

    data_test <-
      data %>%
      mutate(datename = yearmonth(datename))

    data_trn <-
      data %>%
      select(regional_unit, datename, n) %>%
      mutate(datename = yearmonth(datename)) %>%
      fill_gaps() %>%
      slice(1:(n()-cv_dist), .preserve = TRUE) %>%
      stretch_tsibble(.init = init, .step = step)

    data_trn %>% 
      prophetModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)

    p()
  }

  prophet_fits <-
    future_map_dfr(.x =  regions,
                   .f = ~prophetCV(data = filter(data, regional_unit == .x)),
                   .options = furrr_options(seed = TRUE))
  message("prophetModels() complete")

  # Fable models will run using plan(multisession), but need to be run as below
  # with `furrr_options(seed = TRUE)` otherwise improper random numbers are
  # generated
  fable_fits <-
    future_map_dfr(.x =  regions,
                   .f = ~fableCV(data = filter(data, regional_unit == .x)),
                   .options = furrr_options(seed = TRUE))
  message("fableModels() complete")

@@

regionalJointFcsts <- function(data, h ="5 months"){
  regions <- unique(data$regional_unit)

  fable_fit <-
    future_map_dfr(.x = regions,
                   .f = ~fableModels(filter(data, regional_unit ==.x)) %>%
                    forecast(h = h),
                   .options = furrr_options(seed = TRUE)) %>%
    as_tibble()

  message("fable forecasts complete")

  prophet_fit <-
    future_map_dfr(.x = regions,
                   .f = ~prophetModels(filter(data, regional_unit ==.x)) %>%
                    forecast(h = h),
                   .options = furrr_options(seed = TRUE)) %>%
    as_tibble()

  message("prophet forecasts complete")

@@

# Running regionalCV() on each of the quarter of the data
plan(cluster)
system.time(oac_p_gp_cv_models_1 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_1 %>%
              regionalCV())
with_progress(system.time(oac_p_gp_cv_models_2 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_2 %>%
              regionalCV()))
system.time(oac_p_gp_cv_models_3 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_3 %>%
              regionalCV())
system.time(oac_p_gp_cv_models_4 <-
  complete_oac_p_monthly_quantity_gp_pre_cv19_4 %>%
              regionalCV())
plan(sequential)

@@

oac_p_gp_cv_models %>%
  saveRDS("../data/oac_p_gp_cv_models.rds")

@@

```{r}
plan(cluster, workers = 8)
system.time(oac_p_gp_fcsts <-
              oac_p_monthly_quantity_gp_pre_cv19 %>%
              filter(!regional_unit %in% incomplete_gp) %>%
              regionalJointFcsts())
plan(sequential)

oac_p_gp_fcsts %>%
  saveRDS("../data/oac_p_gp_fcsts.rds")

@@

```{r}
oac_p_monthly_quantity_gp_pre_cv19_complete <-
  oac_p_monthly_quantity_gp_pre_cv19 %>% 
  filter(!regional_unit %in% incomplete_gp)

plan(cluster, workers = 8)
system.time(fable_fit <-
  future_map_dfr(.x = oac_p_monthly_quantity_gp_pre_cv19_complete$regional_unit,
                 .f = ~fableModels(filter(oac_p_monthly_quantity_gp_pre_cv19_complete,
                                          regional_unit ==.x)),
                 .options = furrr_options(seed = TRUE)) %>%
  forecast(h = "5 months") %>%
  as_tibble())

system.time(prophet_fit <-
  future_map_dfr(.x = oac_p_monthly_quantity_gp_pre_cv19_complete$regional_unit,
                 .f = ~prophetModels(filter(oac_p_monthly_quantity_gp_pre_cv19_complete,
                                            regional_unit ==.x)),
                 .options = furrr_options(seed = TRUE)) %>%
  forecast(h = "5 months") %>%
  as_tibble())
plan(sequential)

joint_fcst <-
  bind_rows(fable_fcst, prophet_fcst) %>%
  mutate(lb_95 = quantile(n, .025),
         lb_80 = quantile(n, .1),
         med = median(n),
         ub_80 = quantile(n, .9),
         ub_95 = quantile(n, .975),
         sd = distributional::variance(n)^2,
         dist = n,
         n = NULL)
joint_fcst %>%
  saveRDS("../data/oac_p_gp_fcsts.rds")
